---
title: 'DATA 624: PREDICTIVE ANALYTICS HW4'
author: "Gabriel Campos"
date: "Last edited `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_notebook: default
  geometry: left=0.5cm,right=0.5cm,top=1cm,bottom=2cm
  html_document:
    df_print: paged
  pdf_document:
    latex_engine: xelatex
urlcolor: blue
---

```{r}
library(mlbench)
library(dplyr)
library(ggplot2)
library(tsibble)
library(tidyr)
library(corrplot)
library(cowplot)
library(psych)
```


# Instructions

Do problems 3.1 and 3.2 in the Kuhn and Johnson book Applied Predictive Modeling.  Please submit your Rpubs link along with your .pdf for your run code.

# 3.1

The UC Irvine Machine Learning Repository6 contains a data set related
to glass identification. The data consist of 214 glass samples labeled as one
of seven class categories. There are nine predictors, including the refractive
index and percentages of eight elements: Na, Mg, Al, Si, K, Ca, Ba, and Fe.
The data can be accessed via:

```{r}
data(Glass)
str(Glass)
```
## (a)

Using visualizations, explore the predictor variables to understand their
distributions as well as the relationships between predictors.


```{r}
Glass %>%
  select(-10)%>%
  gather() %>%
  ggplot(aes(x=value))+
  geom_histogram(fill="lightgreen")+
  facet_wrap(~key,scales = "free")


corrplot(cor(Glass%>%select(-10)),type="lower")

Glass%>%
  select("RI","Ca")%>%
  ggplot(aes(x=RI,y=Ca))+
  geom_point()+
  geom_smooth(method="lm", se = FALSE,color="red")+
  labs(x = "Refractive Index (RI)", y = "Calcium (Ca)", 
       title = "RI vs Ca Scatterplot" )
Glass%>%
  select("RI","Mg")%>%
  ggplot(aes(x=RI,y=Mg))+
  geom_point()+
  labs(x = "Refractive Index (RI)", y = "Magnesium (Mg)", 
       title = "RI vs Mg Scatterplot" )

```

In order to make the plots `Type` was excluded, as the values were not numeric and did not provide any real insight if it were to be plotted.

Noting the plots I can assess the following:

* All plots with exception of `Mg` and `Si` appear to be right skewed to some degree
* Of the right skewed properties `Ba` and `Fe` is centered around 0
* `Ai` and `Ba` is the most centered
* `RI` and `Ca` is shown to have the highest positive correlation while `RI` and `Mg` has the lowest.
* none of the degrees of relationship seems prominent

## (b)

Do there appear to be any outliers in the data? Are any predictors skewed?

```{r}
Glass %>%
  select(-10)%>%
  gather() %>%
  ggplot(aes(value))+
  geom_boxplot()+
  facet_wrap(~key,scales = "free")
```

All of the attribute except for `Mg` have outliers in the data.
Skewness is shown below. As noted all have a degree of skewness, right and left skewness is described in (a) with `Ba` having the most notable degree.

```{r}
data.frame(describe(Glass))%>%
                      select(skew)
```


## (c)

Are there any relevant transformations of one or more predictors that
might improve the classification model?

# 3.2

The soybean data can also be found at the UC Irvine Machine Learning
Repository. Data were collected to predict disease in 683 soybeans. The 35
predictors are mostly categorical and include information on the environmen￾tal conditions (e.g., temperature, precipitation) and plant conditions (e.g., left
spots, mold growth). The outcome labels consist of 19 distinct classes.

The data can be loaded via:

```{r}
data(Soybean)
## See ?Soybean for details
```

## (a)

Investigate the frequency distributions for the categorical predictors. Are
any of the distributions degenerate in the ways discussed earlier in this
chapter?

## (b)

Roughly 18 % of the data are missing. Are there particular predictors that
are more likely to be missing? Is the pattern of missing data related to
the classes?
## (c) 

Develop a strategy for handling missing data, either by eliminating
predictors or imputation.